---
title: "Exploring titanic Dataset from Kaggle Conpetition""
---

## Getting the Titanic Data
```{r echo=TRUE}
train.data<-read.csv(file="https://raw.githubusercontent.com/agconti/kaggle-titanic/master/data/train.csv",header=T,sep=",")
str(train.data)
```

```{r echo=TRUE}
train.data$Pclass = factor(train.data$Pclass)
train.data$Survived = factor(train.data$Survived)
str(train.data)
```

```{r echo=TRUE}
is.na(train.data$Age)
sum(is.na(train.data$Age) == TRUE)
length(train.data$Age)
sum(is.na(train.data$Name)==T)
length(train.data$Name)
```

# For a title containing a missing value, assign the mean age value for each title not containing a missing value.

```{r echo=TRUE}
mean.mr = mean(train.data$Age[grepl(" Mr\\.",
     train.data$Name) & !is.na(train.data$Age)])
train.data$Age[grepl(" Mr\\.",train.data$Name) & is.na(train.data$Age)]=mean.mr
```

```{r echo=TRUE}
mean.mrs = mean(train.data$Age[grepl(" Mrs\\.", 
      train.data$Name) & !is.na(train.data$Age)])
train.data$Age[grepl(" Mrs\\.",train.data$Name)&is.na(train.data$Age)]=mean.mrs
```
```{r echo=TRUE}
mean.dr = mean(train.data$Age[grepl(" Dr\\.",
      train.data$Name) & !is.na(train.data$Age)])
train.data$Age[grepl(" Dr\\.",train.data$Name) & is.na(train.data$Age)]=mean.dr
```
```{r echo=TRUE}
mean.miss = mean(train.data$Age[grepl(" Miss\\.", 
      train.data$Name) & !is.na(train.data$Age)])
train.data$Age[grepl(" Miss\\.",train.data$Name) & is.na(train.data$Age)]=mean.miss
```
```{r echo=TRUE}
mean.master = mean(train.data$Age[grepl(" Master\\.",
       train.data$Name) & !is.na(train.data$Age)])
train.data$Age[grepl(" Master\\.", train.data$Name) & is.na(train.data$Age)]=mean.master
```

# List the distribution of Port of Embarkation. 
Add the useNA = "always" argument to show the number of NA values
```{r echo=TRUE}
table(train.data$Embarked, useNA = "always")
```

# Assign the two missing values to the most counted port, which is Southampton in this case:
```{r echo=TRUE}
train.data$Embarked[which(is.na(train.data$Embarked))] = 'S'
table(train.data$Embarked, useNA = "always")
```

# Exploring and visualizing data
```{r echo=TRUE}
barplot(table(train.data$Survived), main="Passenger Survival",names= c("Perished", "Survived"))
```
```{r echo=TRUE}
barplot(table(train.data$Pclass), main="Passenger Class", names= c("first","second", "third"))
```
```{r echo=TRUE}
barplot(table(train.data$Sex), main="Passenger Gender")
hist(train.data$Age, main="Passenger Age", xlab = "Age")
```
```{r echo=TRUE}
counts = table( train.data$Survived,train.data$Sex)
barplot(counts, col=c("darkblue","red"),
             legend = c("Perished",  "Survived"), 
               main = "Passenger Survival by Sex")
```
```{r echo=TRUE}
hist(train.data$Age[which(train.data$Survived == "0")], main= "Passenger Age Histogram", xlab="Age", ylab="Count", col ="blue", breaks=seq(0,80,by=2))
hist(train.data$Age[which(train.data$Survived == "1")], col ="red", add=T, breaks=seq(0,80,by=2))
```
